Conceptual and methodological problems with bias detection and avoidance in natural language processing

### 2021-06-15 thesis/03-bayesian.Rmd
Adds up-to-date changes

\todo{Explain where you took the embedding from, cite publications first announcing it.}

@@

\todo{describe which debiasing you used and how it was advertised in the paper}

@@

@@

>>>>

We first encountered the GoogleNews word embedding model in @Bolukbasi2016Man. They use it for their calculations as written in [Tolga Code](https://github.com/tolga-b/debiaswe). We decided to compare the results obtained with the use of Reddit L2 model and the ones that we got by GoogleNews model. The details of the model can be found here [Google News model](https://code.google.com/archive/p/word2vec/). One of the main differences between the models is that Reddit word embeggings have 50 dimensions and GoogleNews word embeddings have 300 dimensions. As the dimension is increasing, the vector can capture much more information but this is not always the case. According to some researchers the amount of dimensions should be chosen by taking into account the type of corpus and its features. In our case both Reddit and GoogleNews models are already trained and ready to use so we do not analyse further the choice of their hyperparameters.

@@

We also applied to the word embeddings hard debiasing method found in @manzini2019black. After debiasing the Reddit word embeddings we calculated the cosine distance again to see if there are any significant changes in terms of the similarity between protected words and the attributes.

@@

One may observe very slight changes in the estimated coeﬀicients. It seems that in religion dataset the changes are the most significant as the mean moves towards zero which stands for no similarity between words.

@@

One may notice that in religion data set

It is worth paying attention to gender debiased dataset. The change is really minor, still zero out of HPDI range. Note how `different` attributes have high similarity value.

At the same time, there is a small improvement in race, at the price of moving neutral terms closer

In @manzini2019black it is mentioned that debiasing with this method that is based on bias component removal may be insufficient. The reason is that words containing similar biases may still cluster together even after debiasing.

### 2021-06-15 thesis/04-discussion.Rmd
Adds up-to-date changes

- 89% highest posterior density intervals

>>>>

- the distributions of cosine distances of whole population and individual words can vary enormously
- including uncertainty shows how in some cases it is not clear why a word is according to this method
classified as anomalous
- debiasing with the method provided by @manzini2019black does not solve the bias issues

What we propose and how it can help:
-
-

# Summary

Future work:
-

<!-- 
remarks:
- delete papers from github/hide them
-

-->

### 2021-06-15 thesis/_book/_main.tex
Adds up-to-date changes

We first encountered the GoogleNews-vectors-negative300 word embedding model in \protect\hyperlink{ref-Bolukbasi2016Man}{Bolukbasi, Chang, Zou, Saligrama, \& Kalai} (\protect\hyperlink{ref-Bolukbasi2016Man}{2016}). They use it for their calculations as mentioned in \url{https://github.com/tolga-b/debiaswe}. We decided to compare the results obtained with the use of Reddit model and the ones that we got by GoogleNews model. The details of GoogleNews model can be found here \url{https://code.google.com/archive/p/word2vec/}. One of the main differences between the models is that Reddit word embeggings have 50 dimensions and GoogleNews word embeddings have 300 dimensions. As the dimension is increasing, the vector can capture much more information but this is not always the case. According to some researchers the amount of dimensions should be chosen by taking into account the type of corpus and its features. In our case both Reddit and GoogleNews models are already trained and ready to use so we do not analyse further the choice of hyperparameters to train them.

@@

\todo{describe which debiasing you used and how it was advertised in the paper}

@@

@@

@@

\begin{itemize}
\tightlist
\item
  89\% highest posterior density intervals
\end{itemize}

>>>>

We first encountered the GoogleNews word embedding model in \protect\hyperlink{ref-Bolukbasi2016Man}{Bolukbasi, Chang, Zou, Saligrama, \& Kalai} (\protect\hyperlink{ref-Bolukbasi2016Man}{2016}). They use it for their calculations as mentioned in \href{https://github.com/tolga-b/debiaswe}{Tolga Code}. We decided to compare the results obtained with the use of Reddit L2 model and the ones that we got by GoogleNews model. The details of GoogleNews model can be found here \href{https://code.google.com/archive/p/word2vec/}{Google News model}. One of the main differences between the models is that Reddit word embeggings have 50 dimensions and GoogleNews word embeddings have 300 dimensions. As the dimension is increasing, the vector can capture much more information but this is not always the case. According to some researchers the amount of dimensions should be chosen by taking into account the type of corpus and its features. In our case both Reddit and GoogleNews models are already trained and ready to use so we do not analyse further the choice of their hyperparameters.

@@

We also applied to the word embeddings hard debiasing method found in (\protect\hyperlink{ref-manzini2019black}{\textbf{manzini2019black?}}). After debiasing the Reddit word embeddings we calculated the cosine distance again to see if there are any significant changes in terms of the similarity between protected words and the attributes.

@@

One may notice

@@

In (\protect\hyperlink{ref-manzini2019black}{\textbf{manzini2019black?}}) it is mentioned that debiasing with this method that is based on bias component removal may be insufficient. The reason is that words containing similar biases may still cluster together even after debiasing.

@@

\begin{itemize}
\tightlist
\item
  the distributions of cosine distances of whole population and individual words can vary enormously
\item
  including uncertainty shows how in some cases it is not clear why a word is according to this method
  classified as anomalous
\item
  debiasing with the method provided by (\protect\hyperlink{ref-manzini2019black}{\textbf{manzini2019black?}}) does not solve the bias issues
\end{itemize}

\hypertarget{summary}{%
\chapter{Summary}\label{summary}}

\hypertarget{future-work}{%
\section{Future work:}\label{future-work}}
